---
title: "Section 3.2 Notes"
format: html
---

```{r}
library(tidyverse)
library(palmerpenguins)
theme_set(theme_minimal())

penguins_sum <- penguins |> group_by(species, year) |>
  summarise(n_penguins = n()) |>
  mutate(year = factor(year))

ggplot(data = penguins_sum, aes(x = year, fill = species)) +
  geom_col(aes(y = n_penguins)) +
  theme_minimal() +
  scale_fill_viridis_d()
```

Example 1: We prefer the second because it's easier to compare the three groups.

Example 2: We prefer the second one because it includes zero.

Ex 3: Same thing as ex 2, depends on the situation though. The first one is preferable, including zero as a y limit for this type of graph is a lot of wasted space.

Ex 4: Pair one i prefer the colour, it uses the position of points so we can easily see which species has high vs low bill depths. Pair 2 i prefer the facet, the first plot has way too many colours.

```{r}
penguins <- penguins |> mutate(species_island_sex = interaction(species, 
                                                                island,
                                                                sex))
ggplot(data = penguins |>
         filter(!is.na(sex)), aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = species_island_sex)) +
  scale_colour_viridis_d() +
  theme_minimal()


ggplot(data = penguins |> filter(!is.na(sex)), 
                                 aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  facet_wrap(~ species_island_sex) +
  theme_minimal()
```

```{r}
## this is our attempt at a better representation
ggplot(data = penguins |> filter(!is.na(sex)), 
                                 aes(x = bill_length_mm, y = bill_depth_mm, colour = sex)) +
  geom_point() +
  facet_grid(species ~ island) +
  theme_minimal()
```


Ex 5: Pie charts are bad because we're bad at estimating angles versus comparing on a common scale.

Ex 6: We prefer the first one because it's easier to distinguish the different regions from eachother. Although the second one does have a higher data-to-ink ratio, being a lollipop plot.

Ex 7:
```{r}
## sequential data
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = species)) +
  theme_minimal() +
  scale_colour_brewer(type = "seq")

## diverging data - good for political party divide
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = species)) +
  theme_minimal() +
  scale_colour_brewer(type = "div")

## qualitative data - this makes the most sense to use in this penguin example
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = species)) +
  theme_minimal() +
  scale_colour_brewer(type = "qual")
```

## Exercise 1
1. For the number of cases variable, would you use a sequential, diverging, or qualitative fill colour scale?

Sequential (number of cases is clearly ordered: more cases would be a more "pop" colour).

2. For the change in the number of cases variable, would you use a sequential, diverging, or qualitative fill colour scale?

Diverging (change can be negative, positive, or zero).

3. For the political party variable, would you use a sequential, diverging, or qualitative fill colour scale?

Qualitative (only two party options, unordered)

4. For the percent population change variable, would you use a sequential, diverging, or qualitative fill colour scale?

Diverging (change can be negative, positive, or zero - neutral colour)


## Exercise 2
```{r}
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(col = species)) +
  scale_color_viridis_d() ## d is for discrete variable

ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(col = body_mass_g)) +
  scale_color_viridis_c() ## c is for continuous variable

ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(col = body_mass_g)) +
  scale_color_viridis_b() ## b is for binning a continuous variable (common in maps)
```

## Exercise 3
```{r}
library(tidyverse)
library(here)
hpi_df <- read_csv(here("data/hpi-tidy.csv"))
ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing, colour = Region)) +
  geom_point() +
  theme(plot.background = element_rect(fill = "purple"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text = element_text(colour = "red"),
        legend.text = element_text(colour = "blue", size = 10),
        axis.line = element_line(arrow = arrow()))
```




